---
title: "TidyTemplate"
date: 2024-04-26
output: html_document
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(lubridate)
library(countrycode)
library(ggthemes)
library(ggmap)

 theme_set(theme_light())

```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2021-05-04")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

water <- tt$water %>% 
	mutate(report_date = mdy(report_date)) %>% 
	rename(lat = lat_deg,
				 lon = lon_deg,
				 country = country_name) %>% 
	separate(water_tech, c("water_tech", "brand"), sep = " - ",
					 fill = "right") %>% 
	mutate(install_year = ifelse(install_year > 2021, NA_real_, install_year)) %>% 
	filter(!country %in% c("Peru", "Timor-Leste", "Dominican Republic"), !is.na(country)) %>% # Remove Peru and Timor-Leste to clean data as it's supposed to be Africa data only
	filter(between(lat, -35, 37),
				 between(lon, -40, 60))
water %>% 
	view()
```


# Initial Exploration of Data

Initial exploration

```{r initial}

water %>% 
	count(status_id)

water %>% 
	count(water_tech, sort = T)

water %>% 
	count(water_source, sort = T)

water %>% 
	count(water_source, water_tech, sort = T)

# removing years with values that make no sense.
water %>% 
	filter(install_year > 2021)

water %>% 
	filter(install_year > 1980) %>% 
	count(install_year) %>% 
	ggplot(aes(install_year,n)) +
	geom_col()

# lots of missing data for these category.
water %>% 
	count(installer, sort = T)

# pay is a free text column
water %>% 
	count(pay, sort = T)

water %>% 
	count(status_id, status, sort = T) %>% 
	view()

water %>% 
	count(country, sort = T) %>% 
	view()

```

# Testing a map
```{r}
water %>% 
	group_by(country) %>% 
	summarize(lat = mean(lat),
						lon = mean(lon)) %>% 
	ggplot(aes(lon,lat, color = country)) +
	geom_point() +
	geom_text(aes(label = country), vjust = 1, hjust = 1)

country_list <- unique(water$country)

africa_map_data <- map_data("world") %>% 
	as_tibble() %>% 
	mutate(continent = countrycode(region, "country.name", "continent")) %>% 
	filter(continent == "Africa")

water %>% 
	filter(country == "Uganda") %>% 
	sample_n(10000) %>%
	ggplot(aes(lon,lat)) +
	geom_polygon(aes(long, lat, group = group),
							color = "gray",
							fill = "white",
							data = africa_map_data,
							size = .3)+
	borders("world", regions = country_list) +
	geom_point(size = .1, alpha = .25) +
	theme_map()
	
	water %>% 
		count(country, sort = T)

```
	
	Start With: Where are recorded water wells in Uganda
```{r}
	# 
	water_uganda <- water %>% 
		filter(country == "Uganda",
				 between(lat, -2,4),
				 between(lon, 28, 40))

	water_uganda_lumped <- water_uganda %>% 
		mutate(water_source = fct_lump(water_source, 5)) %>% 
		replace_na(list(water_source = "Other")) %>% 
		mutate(fct_reorder(water_source, water_source, length, .desc = T))
	
	water_uganda_lumped$water_source %>% levels()
	
	water_uganda_lumped %>% 
		filter(country == "Uganda",
					 between(lat, -2,4),
					 between(lon, 29, 40)) %>% 
		ggplot(aes(lon,lat)) +
		facet_wrap(~water_source) +
		borders("world", regions = "Uganda") +
		geom_point(size = .1, alpha = .25) +
		theme_map() +
		scale_color_discrete(guide = guide_legend(override.aes = list(size = 2, alpha = 1)))
```

```{r}

```

